### 2020-09-19 SuplementaryInformation.tex
Add files via upload

\documentclass{article}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{showkeys}
\usepackage{ragged2e}
\usepackage{cite}
\usepackage{authblk}
\usepackage{xcolor}
\newif{\ifmover}
\newif{\ifquitar}
\renewcommand{\Affilfont}{\small}
\newcommand{\notaL}[1]{{\color{orange}L: #1}}
\newcommand{\notaVA}[1]{{\color{orange}VA: #1}}
\newcommand{\notaVCG}[1]{{\bf \em \color{orange}VCG: #1}}
\definecolor{Addtext}{RGB}{0,0,255}
\begin{document}

	%\title{Thickness optimization in wide range quasi omnidirectional multilayer
	% structures}
	\title{Suplementary information: Thickness optimization in wide range quasi
		omnidirectional 1-D photonic structures}
	\author[1,2,3]{Victor Castillo-Gallardo\thanks{email: {\tt
				victor1\_1@hotmail.com}}}
	\author[1,2,3]{Luis Eduardo Puente-Díaz}
	\author[4]{D. Ariza-Flores}
	\author[3]{Hector Pérez-Aguilar}
	\author[2]{W. Luis Mochán\thanks{email: {\tt
				mochan@fis.unam.mx}}}
	\author[1]{Vivechana Agarwal\thanks{email: {\tt
				vagarwal@uaem.mx}}}
	\affil[1]{Centro de Investigación en Ingeniería y Ciencias Aplicadas,
		Universidad del Estado de Morelos, Av. Universidad 1001
		Col. Chamilpa, Cuernavaca, Morelos 62209, México.}
	\affil[2]{Instituto de Ciencias Físicas, Universidad Nacional Autónoma
		de México, Av. Universidad S/N, Col. Chamilpa, 62210 Cuernavaca,
		Morelos, México.}
	\affil[3]{Facultad de Ciencias Físico Matemáticas,
		Universidad Michoacana de San Nicolás de Hidalgo, Av. Francisco
		J. Múgica S/N 58030, Morelia, Mich., México.}
	\affil[4]{Conacyt-Universidad Autónoma de San Luis Potosí, Karakorum 1470, Lomas 4ta Secc, San Luis Potosí, S.L.P., 78210, México.}
	%\date{{\small {\today}}}
	\maketitle

\section{Development of the transfer matrix}

If you consider that the system only varies on the $z$ axis (as show in Fig.
\ref{Fig1}(a)) , the transfer matrix is $M\left(z_{2},z_{1}\right) ${}{}a
2x2 matrix that relates the components of the electric field E$_{\parallel }$
and the magnetic field H$_{\parallel }$ parallel to the $xy$ plane, and
normal to the axis of the structure, evaluated at any two points $z_{2}$
and $z_{1}$
\begin{equation*}
\left( 
\begin{array}{c}
E_{\Vert } \\ 
H_{\Vert }%
\end{array}%
\right) _{z_{2}}=M\left( z_{2},z_{1}\right) \left( 
\begin{array}{c}
E_{\Vert } \\ 
H_{\Vert }%
\end{array}%
\right) _{z_{1}}.
\end{equation*}%

Many equivalent formulations have been proposed to obtain $M$ . For example, if you 
have a complete multilayer system and it periodically replicates to form an artificial photonic crystal, then Bloch's theorem can be used to describe the normal 
modes of this crystal. According to Bloch's theorem, the modes of a periodic system 
may be written as a superposition of Bloch waves, each of which acquires a 
\emph{phase factor} as it propagates from one period to the next. Therefore, each 
Bloch wave would obey 
\begin{equation}
\begin{pmatrix}
E_{\Vert }^{\pm } \\ 
H_{\Vert }^{\pm }%
\end{pmatrix}%
_{z_{N}}=M%
\begin{pmatrix}
E_{\Vert }^{\pm } \\ 
H_{\Vert }^{\pm }%
\end{pmatrix}%
_{z_{0}}=e^{\pm iKD}%
\begin{pmatrix}
E_{\Vert }^{\pm } \\ 
H_{\Vert }^{\pm }%
\end{pmatrix}%
_{z_{0}},  \label{Bloch}
\end{equation}%
where $D=z_{N}-z_{0}$ is the period, which corresponds to the actual
thickness of the multilayered system, and $\pm K$ represents a 1D Bloch's
vector corresponding to a wave that propagates along $z$-direction. Thus, $\Lambda _{\pm }=e^{\pm iKD}$ are
the eigenvalues of transfer matrix $M$ and $E_{\Vert }^{\pm }$ and $H_{\Vert
}^{\pm }$ are the corresponding eigenvectors. Notice that exactly $\det M=1$
is used, so that the product of eigenvalues ought to be $\Lambda _{+}\Lambda
_{-}=1$, and the dispersion relation of the Bloch modes may be obtained in
principle from 
\begin{equation}
\cos KD=\frac{1}{2}\text{Tr}\,M  \label{dispersion}
\end{equation}%
where T$\text{r}$ denotes the trace. \begin{figure}[tbph]
	\begin{center}
		\includegraphics[width=\textwidth]
		{SIF1Squeme1DPC.pdf}
	\end{center}
	\caption{(a) Chirped-type photonic structure formed by two alternating
		materials a and b, with refractive indices $n_{a}$ and $n_{b}$, and their
		thickness $d_{a}$ and $d_{b}$ gradually increases according to the condition
		of $1/4$ of the wavelength at wich $j$-th. period will be sintonized. 
		\textbf{E}$_{\protect\alpha }^{\protect\beta }$ and \textbf{H}$_{\protect%
			\alpha }^{\protect\beta }$ are the electric and magnetic fields, and \textbf{%
			k}$_{\protect\alpha }^{\protect\beta }$ is the wavevector, where $\alpha
		=\left\{ i,r,a,b\right\} $ corresponds to the medium where the wave is
		traveling and $\beta =\{-,+\}$ to the direction of propagation of the wave
		along the $\pm z$ axis. (b) The multilayer system can be replaced by a broad
		effective layer within which the electric \textbf{E}$_{\protect\beta }$ and
		magnetic \textbf{H}$_{\protect\beta }$ fields, and a Bloch wave \textbf{K}$_{%
			\protect\beta }$ propagate in the positive direction and another set
		propagate in the opposite direction along the $z$ axis.}
	\label{Fig1}
\end{figure}

Now, considering a finite system made of $P$ periods, with $PD$ thickness
and placed on a substrate. For the extreme case of only $P=1$ period, the
optical properties of this system can be obtained by considering an incoming
wave from the ambient, a wave reflected to the environment, a wave
transmitted to the substrate, and two Bloch waves within the structure, one
moving up and other down, as show in Fig. \ref{Fig1}(b).

Hence, the continuity of $E_{\Vert }$ and $H_{\Vert }$ will produce two
equations at each interface, with which the four unknowns are obtained,
namely $r$, $t$ and the amplitudes of two Bloch's waves. In a finite system, 
$K$ can be complex. Even in the presence of very little dissipation, Bloch's
waves should decay as they propagate. Thus, the eigenvalue of ascending mode 
$\Lambda _{+}=e^{iKD}$ is identified as the one that obeys $|\Lambda +|<1$, $%
\text{Im}K>0$, adding a negligible amount of dissipation if necessary to
avoid cases $|\Lambda +|=1$.

From the eigenvalues $\Lambda _{\pm }$ of the transfer matrix, we may obtain
the corresponding eigenvectors $E_{\Vert }^{\pm }$ and $H_{\Vert }^{\pm }$
from Eq. (\ref{Bloch}), and from them, the corresponding surfaces impedances 
\begin{equation}
Z^{\pm }=-\frac{M_{12}}{M_{11}-\Lambda _{\pm }},  \label{Z+-}
\end{equation}%
where $M_{ij}$ with $i=1,2$ and $j=1,2$ denote the elements of the transfer
matrix $M$. By writing the fields at $z_{0}$ and $z_{N}$ as a superposition
of the upward and downward propagating (or decaying) fields, $E_{\Vert
}^{\pm }=Z^{\pm }H_{\Vert }^{\pm }$, we can relate the fields at $z_{N}$ to
the fields at $z_{0}$ through a \emph{reconstructed} transfer matrix, 
\begin{equation}
\begin{pmatrix}
E_{\Vert } \\ 
H_{\Vert }%
\end{pmatrix}%
_{z_{N}}=\tilde{M}%
\begin{pmatrix}
E_{\Vert } \\ 
H_{\Vert }%
\end{pmatrix}%
_{z_{0}},  \label{tilMEH}
\end{equation}%
where 
\begin{equation}
\tilde{M}=\frac{1}{Z^{+}-Z^{-}}%
\begin{pmatrix}
Z^{+}e^{iKD}-Z^{-}e^{-iKD} & -2iZ^{+}Z^{-}\sin KD \\ 
2i\sin KD & Z^{+}e^{-iKD}-Z^{-}e^{iKD}%
\end{pmatrix}%
.  \label{tilM}
\end{equation}%
Finally, the explicit expressions for the optical coefficients are: 
\begin{equation}
r=\mp \frac{Z_{0}\tilde{M}_{11}+\tilde{M}_{12}-Z_{0}Z_{s}\tilde{M}_{21}-Z_{s}%
	\tilde{M}_{22}}{Z_{0}\tilde{M}_{11}-\tilde{M}_{12}-Z_{0}Z_{s}\tilde{M}%
	_{21}+Z_{s}\tilde{M}_{22}},  \label{rBloch}
\end{equation}%
and 
\begin{equation}
t=\frac{2Z_{\alpha }}{Z_{0}\tilde{M}_{11}-\tilde{M}_{12}-Z_{0}Z_{s}\tilde{M}%
	_{21}+Z_{s}\tilde{M}_{22}},  \label{tBloch}
\end{equation}%
where the upper sign $-$ in Eq. (\ref{rBloch}) and the subscript $\alpha =s$
in Eq. (\ref{tBloch}) has been chosen for the case of TE polarization, while
the lower sign $+$ and the subscript $\alpha =0$ correspond to TM
polarization. As usual, the reflectance is given by $R=|r|^{2}$ and the
transmittance by $T=\beta \left\vert t\right\vert ^{2}$ with $\beta
=Z_{0}/Z_{s}$ for the case of TE polarization and $\beta =Z_{s}/Z_{0}$ for
the case of TM polarization.

\section{Chirped-type Bragg mirrors}

As the porosity contrast of the periods increases, the thickness of the structure 
decreases and the average reflectance increases, as shown in table 1. For example, 
you could have structures with a thickness of 10 $\mu$m and their average 
reflectance greater than 90$\%$.

\begin{table}[tbph]
	\caption{ Design parameters obtained after the optimization of functions for maximum average reflectance and minimum thickness }
	\includegraphics[width=\textwidth]
	{SITableOptimized.pdf}
\end{table}

The mapping of the calculated reflectance of the structures designed with the 
optimized parameters is shown in Fig. \ref{Fig2}. Here it is observed that all of 
them have a PBG in the NIR region of the electromagnetic spectrum, i. e.,  
structures designed from functions $f_{1}$ and $f_{3}$ with a porosity contrast of 
$30/76\%$ have the highest PBG localized from 970 to 1370 nm, these are shown in 
panels (a) and (c). \begin{figure}[tbph]
	\begin{center}
		\includegraphics[width=\textwidth]
		{SIReflectanceVisNIR.pdf}
	\end{center}
	\caption{Calculated reflectance of the structures designed with the optimized parameters of $f_{1}$ (left), $f_{2}$ (center) and $f_{3}$ (right) and with the porosity contrasts of $30/76\%$ (upper panels) and $42/76\%$ (lower panels). The dotted lines limit the area of the PBG.}
	\label{Fig2}
\end{figure}

\newpage

\section{Photonic structure with sub-mirrors stacking}

Before showing the design of the structures formed by the stacking of sub-mirrors, 
the transfer matrix trace of individual mirrors formed by one period (Fig. 
\ref{Fig3}(a)) and six periods (Fig. \ref{Fig3}(b)) was calculated. The reflectance 
of each mirror with its respective periods was also calculated (Figs. \ref{Fig3}(c) 
and \ref{Fig3}(d)). Each mirror was designed to reflect a wavelength in the range of 
400-3000 nm and the illumination range was the same. Additionally, the T$r|\text{M}|$ and R$(\lambda,\theta)$ were calculated for light with TE 
polarization and 45${{}^\circ}$ of incidence. In Fig. 3d it is shown that by using 
mirrors that have 6 periods and designed to reflect at wavelengths greater than 1000 
nm, they also reflect in a certain region of the visible spectrum. Then, to 
reinforce the reflection of the electromagnetic waves in the visible, it is enough 
to place a smaller number of periods. This is another reason for the increasing 
number of periods in the 400 to 800 nm region. \begin{figure}[tbph]
	\begin{center}
		\includegraphics[width=\textwidth]
		{SITrazasyRefSE.pdf}
	\end{center}
	\caption{Calculated transfer matrix trace and reflectance of individual mirrors using ((a) and (c)) one period  and ((b) and (d)) six periods, respectively. Each mirror was tuned to a wavelength between 400 and 3000 nm. The calculations were made in the visible and NIR region and at 45${{}^\circ}$ of incidence.}
	\label{Fig3}
\end{figure}
\newpage

It is common to imagine that the stacking of the sub-mirrors is the following: that 
the PBG of the second mirror begins at the edge of the PBG of the first mirror, the 
third PBG that begins at the edge of the PBG of the second mirror, and so on until 
the desired interval is completed. With this configuration, the reflectance 
decreases at each intersection of the two PBGs. For example, in Figs. \ref{Fig4}(a) 
and \ref{Fig4}(b) show the calculated R$(\lambda, \theta)$ for the structures formed 
by sub-mirrors that have one and six periods, respectively, using non-polarized 
light. In these cases, the calculations do not show a PBG, or quasi-PBG. For this 
reason, the R$_{Ave}(\lambda, \theta)$ was maximized through the percentage of 
overlap of the PBGs of two consecutive sub-mirrors. In Fig. 3 (b) of the main 
article the results obtained at the end of the optimization process are shown. The 
calculated R$(\lambda, \theta)$ for the structures designed with 4 periods for each 
sub-mirror and using the pattern described in the main article are shown in Figs. 
\ref{Fig4}(c) and \ref{Fig4}(d) of this document, respectively. Here the overlap of 
the two consecutive PBGs was optimized, being $72\%$ for the first case and $78\%$ 
for the last. Furthermore, the calculations indicate that the structures have a 
quasi-PBG.

\begin{figure}[tbph]
	\begin{center}
		\includegraphics[width=\textwidth]
		{SIReflectanceSEoptimized.pdf}
	\end{center}
	\caption{Reflectance calculated for structures designed through the stacking of 
		sub-mirrors analyzing the overlap of the PBG and the number of periods of 
		each one of them. Case I: there is no overlap of the PBGs and each 
		sub-mirror has (a) one period and (b) six periods. Case II: optimized 
		overlap of consecutive PBGs and (c) each subsmirror has 4 periods with 
		$72\%$ overlap, and (d) proposed arrangement in the main article with 6 
		periods for $\lambda>800$ nm and $78\%$ overlap. The quasi-OD region is bounded by the dotted lines.
	}
	\label{Fig4}
\end{figure}

\end{document}