### 2021-02-18 PaperVictor.tex
Version aprovada por Dra. Vivechana

% !TeX spellcheck = en_US
\documentclass[a4paper,fleqn]{cas-sc}
\usepackage[numbers]{natbib}

%%%Author macros
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
\tsc{EP}
\tsc{PMS}
\tsc{BEC}
\tsc{DE}
%%%
\usepackage{braket}
\usepackage{bm}

\begin{document}
	\let\WriteBookmarks\relax
	\def\floatpagepagefraction{1}
	\def\textpagefraction{.001}
	\shorttitle{Optimization of wide-band quasi-omnidirectional 1-D photonic structures}
	\shortauthors{V. Castillo-Gallardo et~al.}

	\title [mode = title]{Optimization of wide-band quasi-omnidirectional 1-D photonic structures}

	\author[1,2,3]{V. Castillo-Gallardo}
	\cormark[1]
	\ead{ing.victorcg@gmail.com}
	\credit{Conceptualization, Software $\&$ Investigation, synthesis $\&$ characterization, Writing}
	\address[1]{Centro de Investigaci\'{o}n en Ingenier\'{i}a y
		Ciencias Aplicadas, Universidad del Estado de Morelos, Av.
		Universidad 1001, Col. Chamilpa, Cuernavaca, Morelos 62209,
		M\'{e}xico.}

	\author[1,2,3]{Luis Eduardo Puente-D\'{i}az}
	\ead{fmatpuente@gmail.com}
	\credit{Software $\&$ Investigation}
	\address[2]{Instituto de Ciencias F\'{i}sicas, Universidad
		Nacional Aut\'{o}noma de M\'{e}xico, Av. Universidad S/N,
		Col. Chamilpa, 62210 Cuernavaca, Morelos, M\'{e}xico.}
	\address[3]{Facultad de Ciencias F\'{i}sico Matem\'{a}ticas,
		Universidad Michoacana de San Nicol\'{a}s de Hidalgo, Av.
		Francisco J. Múgica S/N 58030, Morelia, Mich., M\'{e}xico.}

	\author[4]{D. Ariza-Flores}
	\ead{david1cool@gmail.com}
	\credit{Investigation, Writing - review $\&$ editing}
	\address[4]{CONACyT-Universidad Aut\'{o}noma de San Luis
		Potos\'{i}, Karakorum 1470, Lomas 4ta Secc, San Luis Potos\'{i},
		S.L.P., 78210, M\'{e}xico.}

	\author[3]{H\'{e}ctor P\'{e}rez-Aguilar}
	\ead{hiperezag@yahoo.com}
	\credit{Investigation and Supervision}

	\author[2]{W. Luis Moch\'{a}n}
	\ead{mochan@fis.unam.mx}
	\credit{Conceptualization, Funding acquisition, Investigation, Supervision, Writing - review $\&$ editing}

	\author[1]{V Agarwal}
	\cormark[2]
	\fnmark[1]
	\ead{vagarwal@uaem.mx}
	\credit{Conceptualization, Funding acquisition, Investigation, Methodology, Supervision, Writing - review $\&$ editing}

	\cortext[cor1]{Corresponding author}
	\cortext[cor2]{Principal corresponding author}
	\fntext[fn1]{On sabatical leave at ICF-UNAM from CIICAP-UAEM}

	\begin{abstract}
		Porous silicon is a very versatile material for developing optical
		devices based on multilayered structures, since high and low porosity layers
		can be synthesized by a simple electrochemical technique. In the present work, we design
		mirrors using \textit{chirped}-type structures and by stacking
		appropriately tuned sub-mirrors, and we optimize the design parameters
		to maximize the
		optical reflectance $\braket{R}$ averaged within a wide
		spectral range  covering the visible and near infrared regions, and
		within a wide angular range. The chirped structures are found to be
		more suitable for the visible range,
		while stacked sub-mirrors are found to be better for the NIR
		region. We design, fabricate and characterize
		some of the optimized omnidirectional mirrors with less
		than 100 pairs of layers. We obtained structures with a wider
		omnidirectional spectral region than those previously
		reported.
	\end{abstract}

	\begin{graphicalabstract}
		\includegraphics[width=\textwidth]{GraphicalAbstract.pdf}
	\end{graphicalabstract}

	\begin{highlights}
		\item We design, optimize, fabricate and characterize wideband quasi-omnidirectional (q-OD) structures using different approaches.
		\item Reflectance averaged over wide spectral range is optimized with simple parametrized functions (chirped) and by stacking Bragg submirrors.
		\item Chirped structure with q-OD bandwidth of 1800 nm in the NIR have been demonstrated.
		\item Stacking Bragg submirrors resulted in similar q-OD band with two third of total physical thickness.
		\item Agreement between the theoretical and experimental results is obtained for porous silicon multilayers .
		\item Proposed design techniques can be useful for obtaining OD dielectric mirrors with different materials in large spectral regions.
	\end{highlights}

	\begin{keywords}
		Optimized reflectance \sep Omnidirectional structures \sep Chirped structures \sep Stacked Bragg mirror structures
	\end{keywords}

	\maketitle

	\section{Introduction}
	Photonic crystal (PC)-based structures have been extensively investigated due
	to their light controlling properties \cite{Joannopoulos2008,Goyal2018}.
	Photons entering a PC interact with its periodically varying dielectric constant and
	their energies are consequently organized into photonic bands. In
	analogy to electronic bands withiny a
	crystal, propagation is forbidden if the energy lies
	within certain regions known as photonic band gaps (PBG's). Thus, the
	PC behaves as a mirror if illuminated by light with frequencies
	within in a PBG. For several applications it is useful to fabricate
	{\em ommnidirectional} mirrors with an absolute band gap for all angles of incidence.
	The use of metallic
	mirrors is limited due to their relatively large absorption at the
	visible (Vis) and near infrared (NIR) frequencies.  An attractive alternative
	consists of low-absorption dielectric
	structures,  which may be designed for specific frequency
	ranges. One-dimensional (1D) and two-dimensional (2D) PCs have
	applications in optoelectronics, optical telecommunications and computing,
	laser technology \cite{Lopez2003,Masaya2010}, and radiative cooling applications
	\cite{Kumar2020}. The simplest 1D-PC is composed of a finite number
	of periodically alternating
	layers of high $n_{H}$ and low $n_{L}$ refractive indices, with
	corresponding thicknesses $d_H$ and $d_L$ chosen so that their optical
	thicknesses corresponds to a quarter of a wavelength
	$n_{H}d_{H}=n_{L}d_{L}=\lambda_{0}/4$ at a nominal
	wavelength $\lambda_{0}$. This yields a {\em Bragg mirror} (BM)
	which might have a large PBG including an omnidirectional gap if the contrast $n_H/n_L$
	and the number of periods are large enough. Mirrors with high
	reflectance within a wider frequency range may be engineered by
	using different nominal wavelengths for different layers.
	One way to obtain such structures is to gradually change the width of
	the layers as a function of their depth, producing
	\textit{chirped}-type structures \cite{Zipock1997}.
	Another alternative is to stack BMs, in such a way that their reflection bands partially
	overlap each other at their edges, resulting in a wider high-reflection band \cite{Xifre2009}.

	On the other hand, different fabrication techniques have been used to
	obtain omnidirectional mirrors (ODM) that operate on the visible and/or
	near infrared (Vis-NIR) range of the electromagnetic spectrum. For
	example, Chen et al. \cite{Chen1999},
	reported the fabrication of a structure made of six pairs of
	TiO$_{2}$/SiO$_{2}$  layers with
	an omnidirectional band (ODB) of about 70 nm in the NIR range, employing a sol-gel deposition
	method. Park et al. \cite{Park2003}, used molecular beam epitaxy to
	grow a stack of four pairs of GaAs/AlAs layers, followed by its conversion to
	GaAs/Al$_{2}$O$_{3}$ by selective oxidation of the AlAs
	layers, to obtain an ODB from 710 to 950 nm.
	On the other hand, DeCorby et al. \cite{DeCorby2005}, fabricated mirrors by
	coupling multiple layers of Ge$_{33}$As$_{12}$Se$_{55}$ chalcogenide glass and
	polyamide-imide deposited by thermal evaporation and spin-casting
	respectively, to obtain a 150 nm wide omnidirectional band centered at
	1750 nm. Furthermore, Jena et al. \cite{Jena2019},
	used sequential asymmetric bipolar pulsed DC magnetron sputtering for TiO$%
	_{2}$ layers and radio frequency magnetron sputtering for SiO$_{2}$
	layers to generate TiO$_{2}$/SiO$_{2}$ 1D-PC's and achieved an ODB from
	592 to 668 nm. However, these techniques are expensive and they
	require sophisticated equipment and long fabrication times. Use of
	porous silicon (PS)
	is an attractive alternative, since it can be easily manufactured by
	electrochemical etching crystalline Si in an hydrofluoric
	acid based electrolyte to obtain a sponge-like nanostructure composed
	of Si and air. By modulating the supplied current and its application
	time during the electrochemical reaction, it is possible to obtain
	multilayered structures allowing for the fabrication ODM's \cite{Xifre2015,Pavesi2000}.
	Although optical filters \cite{Estevez2009,Ariza2014} are the most common
	application of PS PC's, they have also been widely used as chemical
	sensors \cite{Giusseppe2011,Agarwal2018}, waveguides \cite{Hussel1997} and
	for photoluminescence control \cite{Antunez2014}. Recently, the study of
	chirped multilayer structures has
	increased due to their possible application as flat focusing mirrors \cite{Wu2021,Kozar2017,Cheng2018}.
	PS based dielectric optical filters, quasi-ODM's and ODM's have been extensively studied in different
	regions of the electromagnetic spectrum, such as the ultraviolet (UV)
	\cite{Jimenez2020}, visible \cite{Ariza2012}, and NIR \cite{Bruyant2003} regions.
	However, to obtain the desirable high index of refraction
	contrast, layers of high porosity have to be employed, which makes the
	resulting structures relatively fragile. In addition, use of high
	porosity makes the fabrication of large structures, with a
	very large number of layers, unfeasible. For this reason, in this work
	we study different strategies to produce multilayered mirrors
	with high reflectance omnidirectional band
	with relatively small thickness and refractive index contrast.

	This paper is organized as follows. In Sec. \ref{s:theory} we develop the
	methods used to design and calculate the reflectance spectra of the PS
	multilayered structures such as chirped structures and mirror
	stacks. In Sec. \ref{s:exp} we provide details about the
	fabrication of these structures. In Sec. \ref{s:res} we present the numerical
	and experimental results corresponding to our proposed structures and compare
	them to previous reports. We obtained significantly large band-widths
	with relatively thinner structures than reported previously.
	Finally, we discuss our conclusions in Section \ref{s:conc}.

	%\section{Numerical simulation} \label{s:theory}
	\section{Theory} \label{s:theory}

	For the analysis of the propagation of electromagnetic fields through
	multilayered systems it is usual to employ the transfer matrix
	method \cite{PerezA2004,Pochi2005}. Assuming the system is 1D, with variations
	only along the $z$ directions, and that the polarization of light is either
	transverse electric (TE) or transverse magnetic (TM), the
	transfer matrix $M\left(z_{2},z_{1}\right)$ is
	a 2x2 matrix that relates the components of the electric $E_{\| }$
	and magnetic  $H_{\|}$ fields parallel to the $xy$ plane
	evaluated at $z_2$ to their value at $z_{1}$,
	\begin{equation*}
		\left(
		\begin{array}{c}
			E_{\|} \\
			H_{\| }%
		\end{array}%
		\right) _{z_{2}}=M\left( z_{2},z_{1}\right) \left(
		\begin{array}{c}
			E_{\| } \\
			H_{\| }%
		\end{array}%
		\right) _{z_{1}}.
	\end{equation*}%
	Many equivalent formulations have been proposed to obtain and use $M$
	\cite{Mochan1987,Ortiz2020,Chavez2020}. Here we use a recently developed formalism
	for the numerically stable calculation of the reflectance of large
	multilayer systems, as
	proposed by Puente-Díaaz, et al. \cite{Puente2020}, summarized in the
	supplementary information. The
	explicit expressions for the optical coefficients are:
	\begin{equation}
		r=\sigma \frac{Z_{0}M_{11}+M_{12}-Z_{0}Z_{s}M_{21}-Z_s
			M_{22}}{Z_{0}M_{11}-M_{12}-Z_{0}Z_{s}M_{21}+Z_{s}M_{22}},  \label{rBloch}
	\end{equation}%
	and
	\begin{equation}
		t=\frac{2\zeta}{Z_{0}M_{11}-M_{12}-Z_{0}Z_{s}M_{21}+Z_{s}M_{22}},  \label{tBloch}
	\end{equation}%
	where $M_{ij}$ are the elements
	of the transfer matrix which transfer
	the fields from its surface interface with vacuum at $z_0$ towards the
	substrate at $z_s$,  $Z_0$ is the surface impedance of vacuum and $Z_s$ is the
	surface impedance of the substrate, and we define $r=F_r/F_i$ and $t=F_t/F_i$ in terms
	of the reflected and transmitted fields, where we choose the $F$'s as electric
	fields for (TE) polarization and magnetic fields
	for transverse magnetic polarization (TM). Here, $\sigma=-1$ and
	$\zeta= Z_s$ for TE polarization, while
	$\sigma=1$ and $\zeta=Z_0$ for TM
	polarization.
	The reflectance is given by
	$R=|r|^{2}$ and the
	transmittance by $T=\beta \left\vert t\right\vert ^{2}$ with $\beta
	=\text{Re}Z_{0}/\text{Re}Z_{s}$ for TE polarization and $\beta =\text{Re}Z_{s}/\text{Re}Z_{0}$ for
	TM polarization. The complex refractive index of the porous
	layers were obtained through the Bruggeman's effective medium theory, which
	has been reported to adequately reproduce the  optical parameters of PS
	\cite{Giusseppe2011,Pap2006,Estrada2018}.

	\section{Experimental details}\label{s:exp}

	Some of the proposed photonic structures were synthesized through anodic
	etching of a (100) oriented, $p$-type Boron doped, crystalline Si wafer with
	resistivity 0.002-0.005 $\Omega \cdot $cm, under galvanostatic conditions
	\cite{Canham1990,Escorcia2007}. The electrochemical anodization process was
	performed at room temperature, with an electrolyte of aqueous
	hydrofluoric acid (HF) (48$\%$ of
	wt) and ethanol (99.9$\%$ of wt) in 1:1 volumetric proportion, respectively.
	However, as it is not desirable to use very high
	porosity contrasts due to structure fragility and electrolyte
	diffusivity problems \cite{Ariza2011}, the current
	densities were chosen as 35 and 305 $\text{mA}/\text{cm}^{2}$,
	with corresponding porosities of 51\% and 76\%, respectively.
	These porosities were determined through previously obtained calibration
	curves using a gravimetric technique \cite{Pap2006}. The
	etching rate of PS was obtained by synthetizing single layers under
	similar conditions and measuring their thicknesses through Scanning
	Electron Microscopy (SEM). Absolute reflectivity measurements were
	carried out with a Perkin Elmer Lambda 950 UV/Visible
	spectrophotometer with a variable angle universal reflectance
	accessory (URA) for different incident angles $\theta _{i}=10^{\circ
	}$, $20^{\circ }$, $30^{\circ }$, $40^{\circ }$, $50^{\circ }$ and
	$60^{\circ }$ using non-polarized light. The maximum and minimum
	values of $\theta _{i}$ were constrained by the angular range of the
	URA.

	\section{Results and discussion}\label{s:res}

	In this section we present optimized calculations of the reflectance
	of different multilayered omnidirectional wide-band mirrors and we
	compare the results with experimental spectra taken from
	the corresponding samples at several angles of
	incidence. Sec. \ref{ss:chirped} is devoted to chirped-type Bragg
	mirrors while Sec. \ref{ss:stacked} is devoted to structures made of
	stacked sub-mirrors.

	We develop two techniques to design
	multilayer photonic structures. In the first, the wavelength
	($\lambda _{j}$) to which the $j$-th pair of layers is tuned is determined by
	a given function of $j$. Several functions are proposed and their
	parameters are optimized to maximize the reflectance $\braket{R}$ averaged over
	a given spectral and angular range.
	The second technique consisted in stacking
	sub-mirrors tuned to different wavelengths, chosen so that their
	spectral ranges overlap. For each sub-mirror, the spectral range was
	obtained from the dispersion relation of a corresponding periodic
	structure. The degree of overlap was optimized to maximize $\braket{R}$.
	The optimizations were done employing the Minuit module  \cite{minuit}
	of the Perl Data Language (PDL) \cite{PDL1997}.

	\subsection{Chirped-type Bragg mirrors}\label{ss:chirped}

	Here we study multilayered structures where the thicknesses for each
	pair $j=0\ldots N_p-1$ of layers are tuned to a wavelength $\lambda_j$ which changes
	gradually with the depth of the layer, according to
	\begin{equation}
		\lambda_{j}=\lambda_{\min }+\left(\lambda_{\max }-\lambda_{\min}\right) f\left(\frac{j}{N_p-1}\right),\label{Dis}
	\end{equation}%
	where $\lambda_{\min }$ and $\lambda_{\max }$ are the minimum and
	maximum {\em design} wavelengths respectively, $N_p$ is the number of
	periods  in the structure, and $f(\xi)$ is a smooth function that
	goes from 0 to 1 as its argument goes from the surface ($\xi=0$) to the
	substrate ($\xi=1)$. We only consider increasing functions $f$ due to
	the high absorption of PS in the ultraviolet region, which decreases in
	the visible and becomes negligible in the near infrared, so, that the first
	periods are tuned in the UV-Vis regions. We consider the
	following classes of functions:
	\begin{equation}
		f_{1}\left(\xi \right) =\xi^{\alpha },  \label{F1}
	\end{equation}
	\begin{equation}
		f_{2}\left(\xi \right) =\frac{1}{2}\left(\xi^{\alpha }+\xi^{\beta}\right)
		\label{F2}
	\end{equation}
	and%
	\begin{equation}
		f_{3}\left(\xi\right) =A\xi^{\alpha }\left(1-\xi\right) +\xi^{\beta },  \label{F3}
	\end{equation}
	where $\alpha$, $\beta$ and $A$ are parameters to optimize in order
	to maximize the average reflectance $\braket{R}$ over given spectral
	and angular ranges. The class
	$f_{1}$ (Eq. (\ref {F1})) with  $\alpha>0$ corresponds to simple
	increasing profiles, $f_2$ is the arithmetic mean
	of two functions of the type $f_{1}$ with different
	powers $\alpha$ and $\beta$, and $f_{3}$ (Eq. (\ref{F3})) is
	designed so that the first term dominates near the surface and the
	second near the substrate, to allow different behaviors
	at the edges of the spectrum.
	The parameters
	$\alpha$, $\beta$, $A$ and $N_{p}$ were optimized to maximize the
	reflectance for non-polarized light $\braket{R}$ averaged in the
	angular range of $0^\circ$ to $90^\circ$ and either in the
	spectral region $\mathcal R_1$ from 350 to
	1400 nm or $\mathcal R_2$ from 400 to 3000 nm. For $\mathcal R_1$ we took
	$\lambda_{\min}$ as another parameter to be optimized while $\lambda_{\max}$
	was fixed at 1400 nm. For $\mathcal R_2$ we fixed both
	$\lambda_{\min}=850$ nm and $\lambda_{\min}=3000$ nm.
	In the first case, we decided to optimize $\lambda_{min}$ to 
	obtain the best design in the high absorption zone of the PS.
	Approximately $90\%$ of the solar radiation that reaches the
	earth's surface is contained in the spectral region
	$\mathcal R_1$, and $98\%$ in $\mathcal R_2$ \cite{Bird1983}. 
	A direct
	application of the optimized structures in these regions is solar
	reflectors. The
	optimized parameters corresponding to the choice $p_l=0.51$ and
	$p_h=0.76$ for the low and high porosities
	are shown in Table \ref{tab:tableA}.

	\begin{table}
		\centering
		\label{tab:tableA}
		\addtolength{\tabcolsep}{-2pt}
		\begin{tabular}{rrrrrrrrrr}
			Class&$\alpha$&$\beta$&A&$\lambda_{min}(nm)$&$N_{p}$&$d(\mu\text{m})$       &ODW(nm)&ODC (nm)&$\braket{R}(\%)$\\
			\hline
			\hline
			$f_{1}$     &0.24    &---      &---     &250&222   &86.1    &300    &1150    &89\\
			$f_{2}$     &0.37    &1.06     &---     &400&102   &35.9    &220    &1110    &88\\
			$\bm{f_{3}}$&\bf 1.23&\bf 0.540&\bf 0.18&\bf 320&\bf 63&\bf 21.6&\bf 250&\bf 1125&\bf 89\\
			\hline
			$\bm{f_{1}}$&\bf 1.20&\bf ---  &\bf --- &\bf ---&\bf 90&\bf 60.4&\bf 2150&\bf 1925&\bf 92\\
			$f_{2}$     &1.08    &1.86     &---  &---  &105   &61.7    &2000    &2000    &89\\
			$f_{3}$     &1.01    &1.95     &0.20  &---  &102   &62.3    &2050    &1975    &90\\
		\end{tabular}
		\caption{Optimized parameters $\alpha$, $\beta$, $A$ and $N_p$ yielding the highest
			reflectance $\braket{R}$ averaged over the spectral ranges
			$\mathcal R_1$ (first block) and $\mathcal R_2$ (second block) and
			the angular range $0-90^\circ$ for the profile classes $f_1$, $f_2$ and
			$f_3$ (Eqs. (\ref{F1}) - (\ref{F3})). We include the thickness $d$
			of the structure and the omnidirectional spectral
			width and center (ODW and ODC).}
	\end{table}

	The function $f_{3}$ yields after optimization the highest
	$\braket{R}$ in the region $\mathcal R_1$ and with the thinnest
	structure. Furthermore, this
	structure has an omnidirectional width (ODW) of 250 nm
	centered at (ODC) 1125 nm, as shown in Fig. \ref{Fig1}(b).  Various
	photonic structures have been reported with
	ODB's \cite{Xu2018,Ahmed2014,Reiner2000} defined as those spectral regions for
	which $R>90\%$ for all angles. Here we we adopt the same criterium.
	We synthesized a structure designated {\em sample $S_1$} corresponding
	to these results. For the region $\mathcal R_2$ the
	resulting structures have similar thicknesses and attain similar average reflectance,
	though the one with the smallest thickness and the maximum $\braket{R}$
	is that obtained from the family $f_{1}$. We
	also synthesized the corresponding
	structure and designated it {\em sample $S_2$}. The
	calculated optimized reflectance spectrum is shown in Fig. \ref{Fig1}(d).
	This structure has a very wide region where $R(\lambda, \theta)>
	0.90$, which goes from 850 to 3000 nm, but in a restricted angular range
	from $0^\circ$ to $70^\circ$. Thus, it has a
	quasi-ODW of 2150 nm centered at 1925 nm.
	The functions $f_n$ optimized for the region $\mathcal R_1$ are shown
	in Fig. \ref{Fig1}(a), and those for $\mathcal R_2$
	in Fig. \ref{Fig1}(c). The profiles $f_n$ are qualitatively
	similar. For example, for $\mathcal R_1$ the optimal photonic
	structures have a small number of periods tuned to the wavelengths where
	silicon has a large absorption, i.e.,  the first periods of the
	structure.
	As withinin $\mathcal R_2$ the absorption of porous silicon is negligible, the optimized $f_n$ are not too
	far from a linear profile.
	\begin{figure}
		\label{Fig1}
		\begin{center}
			\includegraphics[width=\textwidth]{F1Alternativa.pdf}
		\end{center}
		\caption{Profiles of the functions $f_n(\xi_j)$ optimized for
			the region $\mathcal R_1$ from 350 to 1400 nm (panel a) and
			$\mathcal R_2$ from 400 to 3000 nm (panel c). Reflectance
			for non-polarized light
			vs. wavelength and angle of incidence of the structure
			optimized for the region $\mathcal R_1$ (panel b) and
			$\mathcal R_2$ (panel d). The dotted lines
			indicate the ODB and the dashed lines
			indicate the quasi-ODB.}
	\end{figure}

	In Fig. \ref{Fig2}(a) we present the calculated and measured
	reflectance for non-polarized light corresponding to sample $S_1$ in the spectral range from 350 to 1400
	nm for different angles of
	incidence $\theta=10^\circ$, $20^\circ$, $30^\circ$, $40^\circ$,
	$50^\circ$, and $60^\circ$. The spectra are qualitatively
	similar, and for $\lambda>700$ nm and $10^\circ < \theta <
	40^\circ$, they agree quantitatively, with a difference smaller
	than $5\%$.
	The
	measured reflectance is lower than the calculated one, more so at
	small wavelengths and large angles. This difference
	may be partially attributed to the scattering of light due to the
	roughness present at the sample
	\cite{Theiss1994,Chavez2020,Ortiz2020} that we have not taken into
	account in our theory. Due to the limitations of our
	spectrophotometer, we couldn't perform measurements for $\theta>60^\circ$ and thus
	check the ODB. Nevertheless, our results allow us to conclude that
	the structure has a quasi-ODB from 980 and
	1340 nm.
	\begin{figure}
		\label{Fig2}
		\begin{center}
			\includegraphics[width=\textwidth]{F2Alternativa.pdf}
		\end{center}
		\caption{Reflectance spectra for non-polarized light calculated and measured at different
			angles of incidence $\theta=10^\circ$, $20^\circ$, $30^\circ$,
			$40^\circ$, $50^\circ$, and $60^\circ$ for structure $S_1$
			(panel a)  and $S_2$ (panel b). The gray band indicates the
			region where the reflectance is greater than  $0.90$,
			corresponding to the quasi ODB of the
			structure. Inset: photograph of the synthesized structure.}
	\end{figure}
	In Fig. \ref{Fig2}(b) we compare the
	calculated and measured spectra of sample $S_2$ in the range
	from 400 to 3000 nm.
	Both spectra are very similar, with some differences at long
	wavelengths which can be
	attributed to an increase of imperfections such as wavy layers and
	more roughness for deeper layers, due to a more restricted diffusivity
	of the electrolyte as the thickness of the structure is increased
	\cite{Negro2003,Vincent1993}. According to our results, this
	structure has a wide quasi-ODW of 1800 nm
	centered at 1880 nm.

	\subsection{Photonic structure with sub-mirrors stacking}\label{ss:stacked}

	Another class of techniques for designing good reflectors across a wide spectral range
	is to stack sub-mirrors tuned to different wavelengths chosen with
	different criteria.
	For example, Agarwal et al. \cite{Agarwal2003} proposed a photonic
	structure in which 54 sub-mirrors were stacked as follows: the
	first sub-mirror consisted of 2.5 periods tuned to
	$\lambda_{1}=700$ nm; the following $j$-th sub-mirrors were tuned to
	wavelengths $\lambda_{j}$ according to the sequence $\lambda_{j+1}
	= \lambda_{j}+ \left(2+j \right)$ nm
	with 4 periods for
	each of the following periods.
	Furthermore, Estrada-Wiese et al. \cite{DelRio2018}
	developed a method that combines three stochastic optimization algorithms together with
	a  narrow space search methodology to obtain a customized and optimized
	PC configuration.

	We recall that for a periodic structure, the propagation of
	electromagnetic waves is described by the dispersion relation
	\begin{equation}
		\cos KD = \frac{1}{2} \text{Tr} \bm M,   \label{Dispersion}
	\end{equation}
	where $K$ is Bloch's vector and $\bm M$ the transfer matrix of a single
	period of thickness $D$ \cite{Mochan1987, Mochan1988,
		Perez2018}. Thus,
	whenever $\left\vert\text {Tr} \left(M \right) \right\vert>2$
	electromagnetic energy cannot propagate in the periodic system. This condition corresponds to a
	photonic band gap within which we
	expect a high reflectance, even for a finite system made of only a few periods. This suggests
	the following design strategy. We stack sub-mirrors
	each made of a few periods of a periodic system such that its photonic band gap
	(PBG) overlaps the PBG of the previous and that of the next
	sub-mirror. The number of periods of each sub-mirror is an
	important parameter. If it is too small, the sub-mirror would be
	partially transparent within the PBG. If it is too large, the
	resulting structure would be too thick. This is particularly important
	for sub-mirrors tuned to shorter wavelengths, for which there is some
	absorption. We followed thus two design
	strategies: In the first, we considered sub-mirrors each with the same
	number $p$ of periods. In the second, a different number of periods was
	assigned to each sub-mirror according to the center of its PBG: we chose
	a single period for those sub-mirrors $j$ tuned to
	$\lambda_{j}<500$ nm, two periods for $500 <\lambda_ {j} < 650$ nm,
	three periods for $650<\lambda_ {j}<800$ nm, and for
	$\lambda_{j} \geq 800$ nm the number of periods $p$ was varied between
	4 and 10. In both designs, the degree of overlap of consecutive PBGs was
	taken as the parameter to be optimized in order to maximize
	the reflectance $\braket{R}$ averaged within the spectral and
	angular range $400\text{ nm}\le
	\lambda\le 3000\text{ nm}$ and for $0^\circ\le\theta\le90^\circ$.
	We define the percentage of overlap in terms of the portion of the PBG of the
	$j$-th mirror that lies within the PBG of the $(j + 1)$-th
	mirror.

	Table \ref{tab:table2} shows the optimized values of the average
	reflectance $\braket{R}$ for
	different choices of the number of periods $p$ of each sub-mirror, as
	discussed above, including the resulting number $N_m$ of stacked sub
	mirrors, the thickness $d$ of the structure, and the optimal
	overlap of consecutive PBGs. The first and second blocks correspond to
	the first and second strategies discussed above.
	\begin{table}[htbp]
		\centering
		\label{tab:table2}
		\begin{tabular}{rrrrr}
			$p$&$N_m$&$d (\mu\text{m})$&\% overlap&$\braket{R}$\\
			\hline
			\hline
			1&98&43.2&95&91.0\\
			3&31&42.5&81&91.8\\
			5&17&41.9&68&92.0\\
			\hline
			4&25&38.7&74&92.3\\
			\bf 6&\bf 18&\bf 41.5&\bf 78&\bf 95.4\\
			10&13&45.4&56&95.3\\
		\end{tabular}
		\caption{Total number $N_m$ of stacked sub-mirrors, thickness $d$,
			overlap between consecutive PBG and average reflectance
			$\braket{R}$ for optimized structures designed with a
			number $p$ of periods per sub-mirror (first block) or a number $p$
			of periods for those sub-mirrors tuned to wavelengths
			$\lambda_j\geq 800$ nm (second block, see text), for different
			values of $p$. The parameters corresponding to the best design
			are typeset in bold.}
	\end{table}

	According to table \ref{tab:table2}, the structure that has the highest
	reflectance is in the second block, with an optimal overlap between
	consecutive PBGs of $78\%$, a thickness of
	41.5 $\mu$m, and an average reflectance $\braket{0.954}$. It is made up
	of 18 sub-mirrors tuned to the wavelengths
	400, 460, 510, 570, 640, 720, 810, 910, 1020, 1150, 1290, 1450,
	1630, 1830, 2060, 2320, 2610, and 2860 nm.
	We synthesized this structure and designated it {\em sample $S_3$}.
	Its calculated and measured
	reflectance spectra are shown in Fig. \ref{Fig3}.
	\begin{figure}
		\begin{center}
			\includegraphics[width=\textwidth]
			{F3Alternativa.pdf}
		\end{center}
		\caption{ (a) Calculated reflectance $R(\lambda,\theta)$ for non-polarized light of the optimized structure $S_3$
			as a function of wavelength $\lambda$ and incidence angle
			$\theta$. (b) Measured and calculated reflectance spectra at several
			angles of incidence $\theta=10^\circ$, $20^\circ$,
			$30^\circ$, $40^\circ$, $50^\circ$, $60^\circ$. The gray band indicates
			the region in which the reflectance is greater than $R>0.90$.  Inset shows a
			photograph with a top view of the synthesized structure.}
		\label{Fig3}
	\end{figure}

	Fig. \ref{Fig3}(a) reveals a quasi-ODM
	in the range from 950 to 2900 nm, with an angular range between $0^\circ$ and
	$60^\circ$. It is also observed that at incidence angles $\theta
	>70^\circ$ and long
	wavelengths the reflectance decreases and has oscillations, which we attribute to the
	greater penetration of the electromagnetic field within the structure, as shown by
	Puente-D\'{\i}az, et al. \cite{Puente2020}. In Fig. \ref{Fig3}(b) we compare the
	calculated and measured reflectance for the structure $S_3$ at several angles of
	incidence. Although, the quasi omnidirectional wavelength range of
	the synthesized structure, from 950 to 2750 nm, is slightly smaller than the calculated
	one, the measured PBG is almost twice as wide
	as a recently reported PS chirped structure \cite{Chavez2020}, which
	has the largest previously reported ODW.

	Finally, Table \ref{tab:table3} shows data from some porous
	silicon based ODM designed to operate in the NIR
	region and their of the ODB's are compared with those of
	the structures analyzed in the present study. The ODW
	of the structure $S_1$ results 20 nm wider than that reported by Bruyant,
	et al. \cite{Bruyant2003} and it is located closer to the visible range. Furthermore,
	the ODB of the structure $S_3$ begins at 950 nm like the one
	synthesized by Estevez, et al. \cite{Estevez2009}; nevertheless, its width extends to 2750
	nm. In other words, $S_3$ has an ODW 3.55 times higher. Moreover, the
	ODW of the structure reported by Xifre-Perez, et al. \cite{Xifre2005} represents only 17.6
	\% of the ODW of our $S_2$ and $S_3$ structures. Aditionally, the table
	reveals that the porous silicon based
	structures $S_2$ and $S_3$ have the widest gap widths, 1.36 times greater than the
	largest reported work until now in the NIR region \cite{Wu2021}.

	\begin{table}
		\centering
		\label{tab:table3}
		\begin{tabular}{rrrrr}
			Reference&Year&Sample&ODB (nm)&ODW (nm)\\
			\hline
			\hline
			Bruyant, et al. \cite{Bruyant2003}      &2003&-    &1100-1440&340\\
			Xifre-P\'{e}rez, et al. \cite{Xifre2005}&2004&-    &1297-1615&318\\
			Estevez, et al. \cite{Estevez2009}      &2009&-    &950-1456 &506\\
			Chavez, et al. \cite{Chavez2020}        &2020&-    &1000-2000&1000\\
			Wu, et al. \cite{Wu2021}                &2021&-    &1284-2605&1321\\
			\hline
			&    &$S_1$&980-1340&360 \\
			This work                               &2021&$S_2$&980-2780&1800\\
			&    &$S_3$&950-2750&1800\\
		\end{tabular}
		\caption{Development of porous silicon based omnidirectional mirrors for the NIR region for
			previously published works (first block) and for the different samples studied here (second
			block. We show the attained omnidirectional bands (ODB) and their widths (ODW).}
	\end{table}

	\section{Conclusion}\label{s:conc}

	We have designed, optimized, fabricated and characterized highly reflective quasi
	omnidirectional (angular range of $0-60^\circ$) multilayered structures with a wide
	spectral range. Two techniques, chirping (a continuous change in thicknesses) and
	stacking of Bragg-type sub-structures, have been used to enhance the reflectance with minimum thickness for a given pair of refractive indices. Numerical calculations were
	carried out employing the transfer matrix method and we optimized the
	design parameters to obtain maximal reflectance averaged over different spectral ranges for
	all angles. We fabricated some of the optimized structures with
	porous silicon
	dielectric multilayers and compared their measured  optical properties
	with the calculations. Two chirped structures 
	with thicknesses 21.6 $\mu$m and 60.4 $\mu$m,
	resulting in quasi omnidirectional mirrors with bandwidths of 360 nm and 1800 nm, centered
	at 1160 nm and 1925 nm respectively have been shown. In addition, we fabricated a
	stacked structure  with a thickness of 41.5 $\mu$m resulting in a quasi omnidirectional mirror
	with a bandwidth of 1800 nm centered at 1850 nm.
	Thus, our techniques allowed us to obtain relatively thin quasi-omnidirectional mirrors with wide bands
	over different wavelength ranges. 
	Our analysis techniques can
	be used for the optimization of the reflectance multilayered PS systems with
	different refractive index contrasts or for systems with other types of
	materials. Such proposed structures could be used as mirrors for solar
	concentrators, flat focusing reflectors, thermal regulators, or, if
	defects are included, as filters or remote chemical/biosensors with a
	wide angular independent response.

	\section*{Acknowledgments}

	This project is supported by Consejo Nacional de Ciencia y
	Tecnolog\'{i}a through grant No.A-S1-30393(VA), 256243 (DA), C\'{a}tedras
	Conacyt program 1577 (DA), and by DGAPA-PAPIIT UNAM under grant IN111119 (LM).
	VA acknowledges PRODEP (Sabbatical grant), and
	is grateful to ICF where she spent the sabbatical year.
	HPA also expresses his gratitude to the Coordinaci\'{o}n
	de la Investigaci\'{o}n Cient\'{i}fica de la Universidad
	Michoacana de San Nicol\'{a}s de Hidalgo. VCG is grateful to CIICAp 
	and ICF where he did this work.

	\printcredits

	\bibliographystyle{unsrt}
	\bibliography{Bibliografia}

\end{document}